---
title: "GSEAvalidator"
output: pdf_document
date: "2023-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example workflow

Simple example workflow for how to use GSEAvalidator. Name of protein needs to be 
parsed for genes and is probably where most bugs will happen.

```{r example}
library(GSEAvalidator)
library(tidyverse)

load("../data/example_data.Rdata")

prepped_data = prep_msstats_data(example_data$ProteinLevelData, parse_gene=TRUE)
corr_mat = gen_correlation_matrix(prepped_data, methods = c("pearson"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.3)
test_df
```

## Dataset analysis

Moving forward, no data is included in the package for the GSEA analyses. Please 
contact `kohler.d@northeastern.edu` for access to the data.

### Brunner 2022

#### Hallmark gene sets

``` {r brunner}
load("../data/brunner.Rdata")

prepped_data = prep_msstats_data(brunner$ProteinLevelData, parse_gene=TRUE)
corr_mat = gen_correlation_matrix(prepped_data, 
                                  methods = c("pearson", "spearman"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.5)
test_df

```

#### Regulatory gene sets

``` {r brunner_regulatory}

reg_test_df = test_GSEA_pathways(corr_mat, '../data/regulatory_pathways.json', 
                                 threshold=.5)
reg_test_df

```

### Organoid Data (unpublished)

#### Hallmark gene sets

``` {r org}
load("../data/organoid.Rdata")
organoid_mapper = read.csv("../data/organoid_gene_list.tsv", sep="\t")
prepped_data = prep_msstats_data(organoid$ProteinLevelData, gene_map=organoid_mapper)
corr_mat = gen_correlation_matrix(prepped_data, 
                                  methods = c("pearson", "spearman"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.5)

```

#### Regulatory gene sets

``` {r org_regulatory}

reg_test_df = test_GSEA_pathways(corr_mat, '../data/regulatory_pathways.json', 
                                 threshold=.5)
reg_test_df %>% filter(percent > .5 & percent_measured > .15)

```

### Leduc 2022

#### Hallmark gene sets

``` {r leduc}
leduc = read.csv("/Users/kohler.d/Library/CloudStorage/OneDrive-NortheasternUniversity/Northeastern/Research/MS_data/Single_cell/Leduc/MSstats/MSstats_summarized.csv")
leduc$originalRUN = paste(leduc$Run, leduc$Channel, sep="_")
leduc$LogIntensities = leduc$Abundance
leduc$Protein = unlist(lapply(leduc$Protein, function(x){
  str_split(str_split(x, "\\|")[[1]][[3]], "_")[[1]][[1]]
}))

leduc = leduc %>% dplyr::group_by(originalRUN, Protein) %>% 
  dplyr::summarise(LogIntensities = max(LogIntensities))

prepped_data = prep_msstats_data(leduc)
corr_mat = gen_correlation_matrix(prepped_data, 
                                  methods = c("pearson"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.5)

test_df %>% filter(percent_measured > .2 & 
                         percent > .33)


genes_in_path = extract_genes_in_path(corr_mat, "HALLMARK_ADIPOGENESIS", 
                                      '../data/hallmark_pathways.json')

for (col1 in colnames(prepped_data[, genes_in_path])) {
  for (col2 in colnames(prepped_data[, genes_in_path])) {
    if (col1 != col2) {
      plt = ggplot(prepped_data, aes_string(x = col1, y = col2)) +
        geom_point() +
        geom_smooth(method=lm, se=TRUE, fullrange=TRUE) + 
        labs(title=paste0("Scatter Plot of ", col1, " vs ", col2))
      suppressWarnings(print(plt))
    }
  }
}

```

#### Regulatory gene sets

``` {r leduc_regulatory}

reg_test_df = test_GSEA_pathways(corr_mat, '../data/regulatory_pathways.json', 
                                 threshold=.5)
reg_test_df %>% filter(percent_measured > .2 & percent > .33)

genes_in_path = extract_genes_in_path(corr_mat, "SOX11_TARGET_GENES", 
                                      '../data/regulatory_pathways.json')

for (col1 in colnames(prepped_data[, genes_in_path])) {
  for (col2 in colnames(prepped_data[, genes_in_path])) {
    if (col1 != col2) {
      plt = ggplot(prepped_data, aes_string(x = col1, y = col2)) +
        geom_point() +
          geom_smooth(method = "lm", se = TRUE) + 
        labs(title=paste0("Scatter Plot of ", col1, " vs ", col2))
      print(plt)
    }
  }
}

```

### Montalvo 2022

#### Hallmark gene sets

``` {r montalvo}
load("../data/montalvo.Rdata")
montalvo = montalvo$ProteinLevelData

montalvo$originalRUN = paste(montalvo$Run, montalvo$Channel, sep="_")
montalvo$LogIntensities = montalvo$Abundance
montalvo$Protein = unlist(lapply(montalvo$Protein, function(x){
  str_split(str_split(x, "\\|")[[1]][[3]], "_")[[1]][[1]]
}))

montalvo = montalvo %>% dplyr::group_by(originalRUN, Protein) %>% 
  dplyr::summarise(LogIntensities = max(LogIntensities))


prepped_data = prep_msstats_data(montalvo)
corr_mat = gen_correlation_matrix(prepped_data, 
                                  methods = c("pearson"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.5)
test_df %>% filter(percent_measured > .2 & 
                         percent > .2)

```

#### Regulatory gene sets

``` {r brunner_regulatory}

reg_test_df = test_GSEA_pathways(corr_mat, '../data/regulatory_pathways.json', 
                                 threshold=.5)
reg_test_df %>% filter(percent > .5 & percent_measured > .15)

```
