---
title: "GSEAvalidator"
output: pdf_document
date: "2023-08-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Example workflow

Simple example workflow for how to use GSEAvalidator. Name of protein need to be 
parsed for genes and is probably where most bugs will happen.

```{r example}
library(GSEAvalidator)

load("../data/example_data.Rdata")

prepped_data = prep_msstats_data(example_data$ProteinLevelData, parse_gene=TRUE)
corr_mat = gen_correlation_matrix(prepped_data, methods = c("pearson"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.3)

```

## Dataset analysis

Moving forward, no data is included in the package for the GSEA analyses. Please 
contact `kohler.d@northeastern.edu` for access to the data.

### Brunner 2022

#### Hallmark gene sets

``` {r brunner}
load("../data/brunner.Rdata")

prepped_data = prep_msstats_data(brunner$ProteinLevelData, parse_gene=TRUE)
corr_mat = gen_correlation_matrix(prepped_data, 
                                  methods = c("pearson", "spearman"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.5)
test_df

```

#### Regulatory gene sets

``` {r brunner_regulatory}

reg_test_df = test_GSEA_pathways(corr_mat, '../data/regulatory_pathways.json', 
                                 threshold=.5)
reg_test_df

```

### Organoid Data (unpublished)

#### Hallmark gene sets

``` {r brunner}
load("../data/organoid.Rdata")
organoid_mapper = read.csv("../data/organoid_gene_list.tsv", sep="\t")
prepped_data = prep_msstats_data(organoid$ProteinLevelData, gene_map=organoid_mapper)
corr_mat = gen_correlation_matrix(prepped_data, 
                                  methods = c("pearson", "spearman"))
test_df = test_GSEA_pathways(corr_mat, '../data/hallmark_pathways.json', 
                             threshold=.5)

```

#### Regulatory gene sets

``` {r brunner_regulatory}

reg_test_df = test_GSEA_pathways(corr_mat, '../data/regulatory_pathways.json', 
                                 threshold=.5)
reg_test_df %>% filter(percent > .35)

```

